The Sloan Digital Sky Survey Spectroscopic Reduction Pipeline

### 2005-02-05 doc/paper/paper2d.tex
Update.

The spectra are converted from raw counts (ADUs) to physical
units ($\ergs$).

$ 100 \ergscmang$

@@

We select relatively bright F dwarfs as spectrophotometric 
standards.  Ideally, we want the low-metallicity, halo stars in order 
to have spectra that have few spectroscopic features.
The F stars are chosen to satisfy the following color criteria:
$$ 0.6 < u-g < 1.2 $$
$$ 0.0 < g-r < 0.6 $$
$$ g-r > 0.75 * (u-g) - 0.45 $$
This selects stars that are on the blue edge of the main
sequence in (u-g) color.  At high Galactic latitudes,
??? stars per square degree fall in this color box.
Targetting priority is given to
those stars with colors closest to the following:
$$ u-g = 0.80 $$
$$ g-r = 0.30 $$
$$ r-i = 0.10 $$
This is meant to select those stars closest in color to BD +17 4708,
which is computed to have a similar color (0.93, 0.28, 0.01, courtesy
of David Hogg, priv. comm.).
From these stars, we select 8 on each plate in the magnitude
range $15.5 < g < 17.0$ (which we call SPECTROPHOTO_STD stars)
and 8 in the magnitude range $17.0 < g < 18.5$ (which we call
REDDEN_STD stars).
These magnitude limits are chosen to avoid nearing saturation
of the CCDs on the bright end, but to still ensure good S/N
on the faint end.
Assuming an absolute magnitude of 4.3 for these stars, this
implies distances of $1.7 < d < 3.5$ kpc for the SPECTROPHOTO_STD
stars and $3.5 < d < 6.9$ kpc for the REDDEN_STD stars.
Note that the color-selection and magnitude cuts are done on
``instrinsic'' magnitudes, meaning that we first apply corrections
for SFD extinction.

@@

%\begin{thebibliography}{DUM}
%
%\bibitem[Schlegel, Finkbeiner \& Davis 1998 (SFD)]{sfd98} 
%Schlegel, D. J., Finkbeiner, D. P., \& Davis M. 1998, \apj, 500, 525 (SFD)
%
%\bibitem[York et al. 2000]{york00}
%York, D. G. \etal\ 2000, \aj, 120, 1579
%
%\end{thebibliography}
\clearpage

>>>>

% The spectra are converted from raw counts (ADUs) to physical
% units ($\ergs$).
% $ 100 \ergscmang$

@@

We select relatively bright F dwarfs as spectrophotometric 
standards.  Ideally, we want the low-metallicity, halo stars in order 
to have spectra that have atmospheric opacities that are easy to model.
The F stars are chosen to satisfy the following color criteria:
$$ 0.6 < u-g < 1.2 $$
$$ 0.0 < g-r < 0.6 $$
$$ g-r > 0.75 * (u-g) - 0.45 $$
as shown in Figure ???.
This selects stars that are on the blue edge of the main
sequence in (u-g) color.  At high Galactic latitudes,
??? stars per square degree fall in this color box.
Targetting priority is given to
those stars with colors closest to the following:
$$ u-g = 0.80 $$
$$ g-r = 0.30 $$
$$ r-i = 0.10 $$
This is meant to select those stars closest in color to BD +17 4708,
which is computed to have a similar color (0.93, 0.28, 0.01, courtesy
of David Hogg, priv. comm.).
From these stars, we select 8 on each plate in the magnitude
range $15.5 < g < 17.0$ (which we call SPECTROPHOTO_STD stars)
and 8 in the magnitude range $17.0 < g < 18.5$ (which we call
REDDEN_STD stars).
These magnitude limits are chosen to avoid nearing saturation
of the CCDs on the bright end, but to still ensure good S/N
on the faint end.
Assuming an absolute magnitude of 4.3 for these stars, this
implies distances of $1.7 < d < 3.5$ kpc for the SPECTROPHOTO_STD
stars and $3.5 < d < 6.9$ kpc for the REDDEN_STD stars.
Note that the color-selection and magnitude cuts are done on
``instrinsic'' magnitudes, meaning that we first apply corrections
for SFD extinction under the assumption that these stars are
completely behind the Galactic dust.

The spectra for these stars are compared to the model spectra
from the SPECTRUM code of Gray \& Corbally (1994) as updated
in Gray \etal\ (2001).
We constructed a grid of 90 models with effective temperatures
from $5000$ to $7000 \K$ (spaced every $250 \K$), [Fe/H] values
from -2 to 0 (spaced every 0.5), and gravities of $\log g=4$
and $\log g=4.5$.  These models are then ``observed'' as they
would be seen in our spectra.  We convolve the models with
the instrumental dispersion as a function of wavelength
(see Section ???), and sample the results at the observed wavelengths.
Since there are typically three exposures in the two spectrographs
(blue and red) for each star, this means we re-generate the full
grid of models for six observations of each star.
We perform $\chi^2$ minimizations between these models and the
6 spectra simultaneously.  We use the observed errors per pixel,
which are independent because we have done no resampling of the data.
All spectra and models are first median-filtered with a scale of $7000\kms$,
and the regions of atmospheric absorption in the red are masked
for the purpose of spectral typing our stars.
A velocity in the range $-700 < v < +700 \kms$ is fit
by comparing to the model with $T_{eff} = 6000\K$, Fe/H=$-1.5$,
$\log g=4$.  All the models are redshifted with this best-fit velocity,
and the resulting model with the smallest $\chi^2$ is chosen
as the best-fit model.  We reject any stars with a reduced $\chi^2$
larger than 2.0 in the full spectrum, with a reduced $\chi^2$
larger than 2.0 in the wavelength regions of the strong stellar
absorption lines, or with a median S/N less than 2.0 in
those absorption line regions.
We also reject stars if more than $20\%$ of the pixels are bad
in any of the 3 or more observations, which typically only happens
if a fiber is broken or the spectrum falls on bad columns on the CCD.
Figure ??? shows the distribution of parameters that we have fit
for the ??? flux-calibration stars on the Data Release 4 plates.

Each flux-calibration star provides a means of converting our
observed spectra to physical units.  
In practice, a single star is not sufficient because of the
slight undersampling in wavelength, bad pixels in any
particular observation, stellar typing errors,
and the desire to achieve better S/N in the flux-calibration
vector than one star provides.

Models are redshifted, fluxed, reddened.

@@

\begin{thebibliography}{DUM}

%Gray's code SPECTRUM is
%available http://www.acs.appstate.edu/dept/physics/spectrum/spectrum.html
\bibitem[Gray et al. 2001]{gray01}
Gray, R.\ O., Graham, P.\ W., \& Hoyt, S.\ R. 2001, \aj, 121, 2159

\bibitem[Gray \& Corbally 1994]{gray94}
Gray, R.\ O., \& Corbally, C.\ J. 1994, \aj, 107, 742

\bibitem[Schlegel, Finkbeiner \& Davis 1998 (SFD)]{sfd98} 
Schlegel, D.\ J., Finkbeiner, D.\ P., \& Davis M.\ 1998, \apj, 500, 525 (SFD)

\bibitem[York et al. 2000]{york00}
York, D.\ G. \etal\ 2000, \aj, 120, 1579

\end{thebibliography}
\clearpage
